### 2018-11-11 Sabanci_talk/Sabanci_talk.tex
More slides

\begin{frame}{Neutron Stars}{Most extreme stars in the universe}  
  \begin{small}
    \begin{itemize}
      \item Core-collapse supernova remnants
      \item Mass: $\sim M_\odot$ to $3 M_\odot$
      \item Very \red{compact}: radius $\sim 9$ to 15\,km
      \onslide<2->\item High spins: up to 700 times/second
      \onslide<3->\item Extremely dense: teaspoon $\sim 10^9\,$tons
      \onslide<3->{\item Large magnetic fields: $10^4$ to $10^{11}\,$T}
      \item Hot: surface temperature $10^6\,$K
      \onslide<4->\item Equation of state \\
      Piecewise polytropes
      \[ p = K_i \rho^{\Gamma_i},\quad (\rho_{i-1}\le \rho \le \rho_i) \hspace{8cm}\]
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1>
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.2cm,1.5cm)}]current page.center) {\includegraphics[height=3.2cm]{figs/Core_collapseSXS.jpg}};    
	\node[anchor=center,scale=1,opacity=1] at ([shift={(5.4cm,2.9cm)}]current page.center) {\textcolor{white}{SXS}};
      \onslide<2>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=6cm]{figs/Millisecond_NS.png}}; 
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.8cm,-1.cm)}]current page.center) {Class.Quant.Grav. 30 (2013) 224003};
      \onslide<3>\node (img2 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.6cm,0.0cm)}]current page.center) {\includegraphics[height=4cm]{figs/NS_cross_section.png}}; 
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.6cm,-1.5cm)}]current page.center) {Robert Schulze};
       \onslide<4>\node (img3 )[anchor=center,scale=0.8,opacity=1] at ([shift={(2.7cm,-2.5cm)}]current page.center) {\includegraphics[height=5cm]{figs/NS_EOS_Lattimer2012.png}}; 
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.cm,-3.8cm)}]current page.center) {Lattimer arXiv:1305.3510};
    \end{tikzpicture}
\end{frame}

@@

\begin{frame}{GW Astronomy in the 2020s}{Advanced LIGO-Virgo}
  Advance \red{warning time}: inspiral time from $\bar{f}$
  \[
    T_\text{AW} = \tau_\text{insp}(\bar{f})
  \]
  \onslide<2->
  {
    Total accumulated SNR over the inspiral
    \begin{footnotesize}
      \[
      \bar{\rho}_\text{F}\equiv\sqrt{\left[{\rho}_h(f_0^h,{f}_\text{ISCO})\right]^2+\left[{\rho}_m(f_0^m,{f}_\text{ISCO})\right]^2+\left[{\rho}_l(f_0^l,{f}_\text{ISCO})\right]^2}
      \]
      \begin{table}[h]
	\centering
	\begin{tabular}{lcccccc}
	  \hline\hline
	  $D\,$(Mpc) & $\bar{f}\,$(Hz) &{}& $T^\text{0PN}_\text{AW}$(sec) &\hspace{1mm} & $T^\text{3.5PN}_\text{AW}\text{ (sec)}\hspace{2mm}$& $\bar{\rho}_F$\T \B \\
	  \hline
	  50 & $\approx\,$25.3 & & 84.25 & & 85.86 & 70.3 \\
	  100 & $\approx\,$41.6 & & 22.39 & & 22.86 & 35.2 \\
	  200 & $\approx\,$121 & & 1.291 &\quad & 1.300 & 17.6 \\
	  \hline\hline
	\end{tabular}
      \end{table}
    \end{footnotesize}
  }
%
\onslide<3->$\Longrightarrow$ NO likely FORECASTS  in the 2020s
\begin{tikzpicture}[overlay,remember picture]
  \onslide<1->\node[draw, red,thick,minimum width=2.8cm,minimum height=1.0cm] (b) at (-0.7,5.3){};
\end{tikzpicture}

@@

\begin{frame}{GW Astronomy in the 2030s}{The era of Einstein Telescope}
 One detector, \red{three V}-IFOs, $60^\circ$ 10\,km arms.\\
 Underground,cryogenic, \bl{xylophone} (high, low-frequency modes)\\
 \vspace{1mm}
 Cosmological range $\Rightarrow$ sky-averaged,\red{red}-shifted strain and SNR
 \begin{footnotesize}
 \begin{align*} 
  \tilde{H}_\text{ET}(f) &= \f{3}{5}A h_0 \red{(1+z)^{-1}} f^{-7/6}\\
  \bar\rho_{\text{ET}}(f_0,f)&\equiv \f{6}{5}A h_0\, \red{(1+z)^{-1/6}}\left[\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')}\right]^{1/2}
 \end{align*} 
 \end{footnotesize}
Consider BNS at \\
$$ D=\{100,200, 400, 600\}\,\text{Mpc} \Longrightarrow z=\{0.022, 0.044, 0.085, 0.12\}$$
$\Omega_\Lambda = 0.6911, \Omega_m = 0.3089, H_0 = 67.74\,$km\,s$^{-1}\,$Mpc$^{-1}$

>>>>

\begin{frame}{Neutron Stars}{Most extreme stars in the universe}  
  \begin{small}
    \begin{itemize}
      \item Core-collapse supernova remnants
      \item Mass: $\sim M_\odot$ to $3 M_\odot$
      \item Very \red{compact}: radius $\sim 9$ to 15\,km
      \onslide<2->\item High spins: up to 700 times/second
      \onslide<3->\item Extremely dense: teaspoon $\sim 10^9\,$tons
      \onslide<3->{\item Large magnetic fields: $10^4$ to $10^{11}\,$T}
      \item Hot: surface temperature $10^6\,$K
      \onslide<4->\item Equation of state \\
      Piecewise polytropes
      \[ p = K_i \rho^{\Gamma_i},\quad (\rho_{i-1}\le \rho \le \rho_i) \hspace{8cm}\]
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1>
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.2cm,1.5cm)}]current page.center) {\includegraphics[height=3.2cm]{figs/Core_collapseSXS.jpg}};    
	\node[anchor=center,scale=1,opacity=1] at ([shift={(5.4cm,2.9cm)}]current page.center) {\textcolor{white}{SXS}};
      \onslide<2>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=6cm]{figs/Millisecond_NS.png}}; 
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.8cm,-1.cm)}]current page.center) {Class.Quant.Grav. 30 (2013) 224003};
      \onslide<3>\node (img2 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.6cm,0.0cm)}]current page.center) {\includegraphics[height=4cm]{figs/NS_cross_section.png}}; 
      % Use Fig 3 of http://science.sciencemag.org/content/304/5670/536/tab-figures-data ?
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.6cm,-1.5cm)}]current page.center) {Robert Schulze};
       \onslide<4>\node (img3 )[anchor=center,scale=0.8,opacity=1] at ([shift={(2.7cm,-2.5cm)}]current page.center) {\includegraphics[height=5cm]{figs/NS_EOS_Lattimer2012.png}}; 
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.cm,-3.8cm)}]current page.center) {Lattimer arXiv:1305.3510};
    \end{tikzpicture}
\end{frame}

@@

\begin{frame}{GW Astronomy in the 2020s}{Advanced LIGO-Virgo}
  Advance \red{warning time}: inspiral time from $\bar{f}$
  \[
    T_\text{AW} = \tau_\text{insp}(\bar{f})
  \]
  \onslide<2->
  {
    Total accumulated SNR over the inspiral
    \begin{footnotesize}
      \[
      \bar{\rho}_\text{F}\equiv\sqrt{\left[{\rho}_h(f_0^h,{f}_\text{ISCO})\right]^2+\left[{\rho}_m(f_0^m,{f}_\text{ISCO})\right]^2+\left[{\rho}_l(f_0^l,{f}_\text{ISCO})\right]^2}
      \]
      \begin{table}[h]
	\centering
	\begin{tabular}{lcccccc}
	  \hline\hline
	  $D\,$(Mpc) & $\bar{f}\,$(Hz) &{}& $T^\text{0PN}_\text{AW}$(sec) &\hspace{1mm} & $T^\text{3.5PN}_\text{AW}\text{ (sec)}\hspace{2mm}$& $\bar{\rho}_F$\T \B \\
	  \hline
	  50 & $\approx\,$25.3 & & 84.25 & & 85.86 & 70.3 \\
	  100 & $\approx\,$41.6 & & 22.39 & & 22.86 & 35.2 \\
	  200 & $\approx\,$121 & & 1.291 &\quad & 1.300 & 17.6 \\
	  \hline\hline
	\end{tabular}
      \end{table}
    \end{footnotesize}
  }
%
\onslide<3->$\Longrightarrow$ NO likely FORECASTS  in the 2020s
\begin{tikzpicture}[overlay,remember picture]
  \onslide<1->\node[draw, red,thick,minimum width=2.8cm,minimum height=1.0cm] (b) at (-0.7,5.3){};
  \onslide<3->\node[draw, red,thick,minimum width=1.1cm,minimum height=1.2cm] (b) at (1.6,1.35){};
\end{tikzpicture}

@@

\begin{frame}{GW Astronomy in the 2030s}{The era of Einstein Telescope}
  \vspace{-1mm}
  \begin{itemize}
    \item[] One detector, \red{three V}-IFOs, $60^\circ$ 10\,km arms.
    \item[] Underground,cryogenic, \bl{xylophone} (high, low-frequency modes)
    \uncover<2->{
    \item[] Recall: \red{RMS}-average over all four angles
    \item[] Cosmological range $\Rightarrow$ \red{red}-shifted strain and SNR}
  \end{itemize}
  \onslide<2->
    \begin{footnotesize}
    \begin{align*} 
      \tilde{H}_\text{ET}(f) &= \f{3}{5}A h_0 \red{(1+z)^{-1}} f^{-7/6}\\
      \bar\rho_{\text{ET}}(f_0,f)&\equiv \f{6}{5}A h_0\, \red{(1+z)^{-1/6}}\left[\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')}\right]^{1/2}
    \end{align*} 
    \end{footnotesize}
    \vspace{-1mm}
  \onslide<3->
    Consider BNS at
    \begin{footnotesize}
      $$ \bl{D}=\{100,200, 400, 600\}\,\text{Mpc} \Longrightarrow \red{z}=\{0.022, 0.044, 0.085, 0.12\}$$
      $$ \left(\text{Using } \Omega_\Lambda = 0.6911, \Omega_m = 0.3089, H_0 = 67.74\,\text{km}\,\text{s}^{-1}\,\text{Mpc}^{-1}\right)\qquad  $$
    \end{footnotesize}
  \onslide<4->
  {{\bf Note}: for binary black holes, ET's range $>\,1$Gpc, but inspirals too quick.}
\end{frame}

\begin{frame}{The era of Einstein Telescope}{BNS detections}
 Same detection criterion:\qquad $\bar{\rho}_\text{ET}(f_0, \bar{f}_\text{ET}) =15$ \todo{ADD LIGO2017}\\
 \red{Two} ET configurations: B (high $f$), C (low $f$)\todo{2020 to plot}
 \begin{tikzpicture}[overlay,remember picture]
      \onslide<1->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.5cm)}]current page.center) {\includegraphics[height=6cm]{../Figures/ET_strains_redshifted.pdf}};
      %\onslide<2->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.5cm)}]current page.center) {\includegraphics[height=6cm]{../Figures/ET_strains_redshifted.pdf}};
      %\onslide<3->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.5cm)}]current page.center) {\includegraphics[height=6cm]{../Figures/ET_strains_redshifted.pdf}};
      %\onslide<4->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(0cm,-1.5cm)}]current page.center) {\includegraphics[height=6cm]{../Figures/ET_strains_redshifted.pdf}};
\end{tikzpicture}      
\end{frame}

\begin{frame}{The era of Einstein Telescope}{GRB \red{Forecast} Prospects} 
  \vspace{-8mm}
  \begin{footnotesize}
  \begin{table}[h]
  \centering
  \begin{tabular}{lccccccc}
  %toprule
  \hline\hline
  $D\,$(Mpc) & \multicolumn{3}{c}{ET-B} & $R_\text{av}{}^\text{max}_\text{min} $ & \multicolumn{3}{c}{ET-C}\T\B\\
  \hline
  {}& $\bar{f}_\text{ET}\,$(Hz) & \ \hspace{3mm} $T_\text{AW}$ \ \hspace{3mm} & $\bar{\rho}_{F}$ & {}  & $\bar{f}_\text{ET}\,$(Hz) & \ \hspace{3mm} $T_\text{AW}\hspace{3mm}$& $\bar{\rho}_{F}$\T\B\\
  %midrule
  100 & $\approx\,$6.72 &  47.0\,minutes & 306 & $ 1.54^{4.7}_{0.3}$ & $\approx\,$3.27 & 5.34\,hours\ & 365\T\\
  200 & $\approx\,$11.2 & 11.6\,minutes & 152 &  $ 12.3^{38}_{2.6} $ & $\approx\,$4.10 & 2.87\,hours\ & 182 \\
  400 & $\approx\,$18.2 & 3.00\,minutes & 75.7 &$ 98.6^{300}_{20} $ & $\approx\,$5.06 & 1.51\,hours\ & 90.5\\
  600 & $\approx\,$?  &?\,seconds & ?& $ 333^{1020}_{69} $ &    $\approx\,$5.7 & 1.03\,hours & 60.?  \\
  \hline\hline
  %bottomrule
  \end{tabular}
  \caption*{Event rates using $R=1540^{+3200}_{-1220}\,\text{Gpc}^{-3}\,\text{yr}^{-1}$ inferred from GW170817}
  \end{table}
  \end{footnotesize}
  %
  \vspace{-3mm}
  \onslide<2->Very \red{high SNRs} $\Rightarrow \ord(\rho_F^{\red{-1}})$-precision parameter estimation afterward \\
  \vspace{2mm}
  \onslide<3->\bl{ET-B} $\Longrightarrow $ forecasts not likely \\
  \vspace{2mm}
  \onslide<4->\red{ET-C} $\Longrightarrow T_\text{AW} \ge 1\,$hour within 600\,Mpc\\
  \vspace{1mm}
  \textcolor{white}{ET-C} $\Longrightarrow \red{\ord(100)}$ {\bf BNS forecasts} yearly! $\Rightarrow $ 1 per day!\\
  \vspace{1mm}
  \onslide<5->\textcolor{white}{ET-C} $\Longrightarrow \red{\ord(10)}$ with \red{3\,hours} of warning \\
  \vspace{1mm}
  \onslide<6->\textcolor{white}{ET-C} $\Longrightarrow \red{\ord(2)}$ with \red{5\,hours} of warning
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<3>\node[draw, blue,thick,minimum width=1.8cm,minimum height=1.6cm] (b) at (-2.3,5.3){};
      \onslide<4->\node[draw, red,thick,minimum width=1.8cm,minimum height=1.6cm] (b) at (3.6,5.3){};
  \end{tikzpicture}
\end{frame}

\begin{frame}{The era of Einstein Telescope}{Source localization before the merger} 
  ET is very sensitive, but all 3 IFOs are at one place. \\
  \vspace{1mm}
  $\Longrightarrow$ \red{poor} sky localization \\
  \vspace{1mm}
  \onslide<2->BUT \ldots \\
  \begin{itemize}
  \item With $T_\text{AW} \sim\,$hours, Earth's rotation matters!
  \item Two hours $\Rightarrow 30^\circ \Rightarrow \approx 2000\,$km of IFO distance $\sim\,$L1-H1 separation
  \end{itemize}
  \onslide<3->So, within 1\,Gpc, ET {\bf alone} will localize
  \begin{itemize}
  \item 20\% of the sources to within 100\,deg$^2$
  \item 2\% to within 20\,deg$^2$
  \item 0.5\% to within 10\,deg$^2$ 
  \onslide<4->\item[]$\Longrightarrow \red{\ord(5)}$ sources within 400\,Mpc
  \end{itemize}
  \onslide<5->10\,deg$^2$: LSST's field of view\\
  \onslide<6-> Numbers improve with \bl{Cosmic Explorer} ($f\gtrsim \red{5}\,$Hz)
  %
  \begin{tikzpicture}[overlay,remember picture]
	\onslide<2->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.2cm,2.25cm)}]current page.center) {\includegraphics[height=3cm]{figs/Earth_rot.png}};
	\onslide<5->\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.3cm,-2.25cm)}]current page.center) {\includegraphics[height=3.9cm]{figs/LSST.jpg}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{The era of Einstein Telescope}{Black hole - neutron star binaries}
